---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


[url link for an inline html preview:](http://htmlpreview.github.io/?https://github.com/chatard/RepData_PeerAssessment1/blob/master/PA1_template.html)

## Loading and preprocessing the data

### Preliminary settings and libraries packages:
#### Local date and time settings for US compatibility 

```{r USlocaltime}
Sys.setlocale("LC_TIME","en_US.UTF-8")
```

#### Knitr options:  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Loading basic packages.

```{r packages}
library(dplyr)
library(ggplot2)
```

### Loading data:  

```{r loading}
        if(!file.exists("activity.csv")){
        unzip("activity.zip")
        }
data<- read.csv("activity.csv")
```


###Data characteristics:

#### brief look:  

```{r data}
str(data)
```

#### dimensions:

```{r dimensions}
dim(data)
```

## What is mean total number of steps taken per day?

To calculate this mean value,  we must know the value $S_{i}$ for each day which is 
equal to the total of steps per day and then sum all Si: $\sum S_{i}$ ... then divide by 
the number of days. 
Or, which is the same thing, calculate mean(S)
which is the mean of the variable S.

```{r}
grdata<- data %>%
        group_by(date)%>%
        summarise(stepsmeanbyday=sum(steps, na.rm = TRUE))
```




###Making a histogram of the total number of steps taken each day:        

```{r histogram1, fig.height= 4}
ggplot(grdata, aes(x = stepsmeanbyday)) + 
        geom_histogram(fill = "blue", binwidth = 1000)+
        labs(title="Histogram of Steps Taken per Day", 
             x="Number of Steps per Day",
             y="Number of times in a day")
```


### Mean and median of the total number of steps taken per day:  
    

```{r daymean}
  grmean<- mean(grdata$stepsmeanbyday, na.rm = TRUE)
  
```
  mean  =   `r toString(grmean) `
  
```{r daymedian}  
grmedian <- median(grdata$stepsmeanbyday, na.rm = TRUE)

```

median =   `r toString(grmedian) `


## What is the average daily activity pattern?  


```{r, average}
dataclone<-data 
dataclone$interval.factor <- factor(dataclone$interval)
```

```{r}
data_interval<-dataclone%>% 
                group_by(interval) %>%
                summarise(stepsmean=mean(steps, na.rm = TRUE))
```


```{r, change2numeric}
data_interval$stepsmean<-as.numeric(data_interval$stepsmean)
```

###time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):   

```{r, dayaveragefig, fig.height= 4}
#ggplot(data_interval, aes(interval, as.numeric(stepsmean))) +
        #geom_line(col="blue")
ggplot(data_interval, aes(interval, as.numeric(stepsmean))) +
        geom_line(col="blue")+
        labs(title="Average Daily activity pattern",
             x="x-axis:five minutes intervals", y="average number of steps")
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  

#### maximum number of steps:

```{r, maxistep  }
maxisteps<-max(data_interval$stepsmean)
```

maximum nb. of steps = `r toString(maxisteps) `

#### 5-minute interval that contains the maximum number of steps:

```{r, maxiinterv}

maxi5interval<-data_interval$interval[which.max(data_interval$stepsmean)]
```

maxi-5 interval = `r toString(maxi5interval)`

## Imputing missing values

###total number of missing values in the dataset:   

```{r, totalna}
totalNA<-sum(is.na(data))
```

**totalNA** = `r toString(totalNA)`  

###strategy for filling in all of the missing values in the dataset:

the simplest strategy would be to replace the missing values by the average number of steps taken in a day by example.  

But, if we take a look at the distribution of the variable that corresponds to the average number of steps per day ... it seems that it is a bimodal distribution : 


```{r density, fig.height= 3}
p <- ggplot(grdata, aes(x=stepsmeanbyday)) + 
        geom_density(color="darkblue", fill="lightblue") +
        labs(x= "steps mean by day", y= "probability (mean steps by day =x)")+
        geom_vline(aes( xintercept = mean(stepsmeanbyday) ),
        linetype=  "solid", color =  "green" ) +
        geom_vline(aes( xintercept =median(stepsmeanbyday) ),
                   linetype=  "solid", color =  "red" )    
```

```{r}
p
```

**So, we could try to refine our strategy by replacing, the weekend missing values by the average of the steps that correspond to the weekend days and to replace the missing values of the ordinary days by the average calculated on the days of weeks**.


## Are there differences in activity patterns between weekdays and weekends?
