---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
    number_sections: true
---

# Peer Assignment 1 :  
[inline html preview:](http://htmlpreview.github.io/?https://github.com/chatard/RepData_PeerAssessment1/blob/master/PA1_template.html)

## Loading and preprocessing the data

### Preliminary settings:
#### Local date and time settings for US compatibility 

```{r USlocaltime}
Sys.setlocale("LC_TIME","en_US.UTF-8")
```

#### Knitr options:  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Loading basic packages.

```{r packages}
library(dplyr)
library(ggplot2)
```

### Loading data:  

```{r loading}
        if(!file.exists("activity.csv")){
        unzip("activity.zip")
        }
data<- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
